### 2021-03-09 draft2.Rmd
rmd file

---
title: "Movie Recommendation System"
author: "Ray Sherbourne"
date: "March 4, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Project Overview
###section that describes the dataset and summarizes the goal of the project and key steps that were performed

The goal of this project is to solve a popular data science problem known as **The Netflix Challenge**. A dataset of movies and user ratings is provided that will serve as the foundation to build a recommendation system. A validation set is extracted from the data to simulate a true unknown sample and to serve as a final validation test set. The accuracy, as measured against the validation set, is specified as the root mean squared error (RMSE), with a target of 0.86499 or better.

The key steps that were performed are:
1. Import data
  + Download data from source
  + Extract validation set
2. Data wrangling to make data more intuitive
3. Exploratory analysis
4. Model building
  + Start with the mean as a baseline approach
  + Introduce a movie bias term
  + Introduce a user bias term
  + Introduce a regularization term lambda
  + Use matrix factorization
5. Train with best performing model and test on validation set
6. Reflect on the results

#Method
###section that explains the process and techniques used, including data cleaning, data exploration and visualization, insights gained, and your modeling approach

The data was downloaded from the grouplens website and 10% was immidiately set aside as the validation set. Looking at the test set, there appears to be two useful data munging operations we can do. The timestamp predictor appears as a comon format of seconds from xxx. Although this is a common format, it is unintuitive to work with so this is converted to a date time format. The release year may also have some value to our exploration so it is extracted from the movie title.

To start building a model, we will first use the simplest approach to use as a benchmark and itteratively improve upon it. The first model will predict the average movie rating for all movies.

We know that not all movies are recieved by audiences equally.

#Results
###section that presents the modeling results and discusses the model performance

#Conclusion
###section that gives a brief summary of the report, its limitations and future work