### 2021-03-12 draft2.Rmd
finished outline of #method, ready to add code chunks

---
title: "Movie Recommendation System"
author: "Ray Sherbourne"
date: "March 4, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Project Overview
###section that describes the dataset and summarizes the goal of the project and key steps that were performed

The goal of this project is to solve a popular data science problem known as **The Netflix Challenge**. A dataset of movies and user ratings is provided that will serve as the foundation to build a recommendation system. A validation set is extracted from the data to simulate a true unknown sample and to serve as a final validation test set. The accuracy, as measured against the validation set, is specified as the root mean squared error (RMSE), with a target of 0.86499 or better.

The key steps that were performed are:
1. Import data
  + Download data from source
  + Extract validation set
2. Data wrangling to make data more intuitive
3. Exploratory analysis
4. Model building
  + Start with the mean as a baseline approach
  + Introduce a movie bias term
  + Introduce a user bias term
  + Introduce a regularization term lambda
  + Use matrix factorization (with and without optimizing parameters)
5. Train with best performing model and test on validation set
6. Reflect on the results

#Method
###section that explains the process and techniques used, including data cleaning, data exploration and visualization, insights gained, and your modeling approach

The data was downloaded from the grouplens website and 10% was immediately set aside as the validation set. Looking at the test set, there appears to be two useful data munging operations we can do immediately. The timestamp predictor appears as a common format of seconds since midnight (UTC) January 1st, 1970. Although this is a common format, it is un-intuitive to work with so this is converted to a date time format. The release year may also have some value to our exploration so it is extracted from the movie title.

Let's now visualize each of our predictors, aswell as some hypothesis that could merrit further investigation.

We will use the simplest approach as a benchmark and itteratively improve upon it. Therefore, our first model will predict the average movie rating for all movies.

To account for the fact that movies are not received by audiences equally, we can introducing a bias term. This first bias term will represent how a particular movie deviates from the average of all movies.

Likewise, different users have different movie critiquing tendencies. We can introduce a bias term for users that will represent how a particular user deviates from the average user in rating movies.

Now, let's see how the results would change if we reversed the order of the calculation of our bias terms.

Taking the best form of our biased model, we can introduce a regularization term, lambda, to combat over-fitting. In the context of the Netflix challenge, over-fitting could occur due to a rarely watched movie being incorrectly labeled as exceptionally good and '`leading to unhelpful recommendations. We will use cross-validation to search for the optimal lambda term.

We will now try an alternate approach using matrix factorization. The idea behind this approach is that if we arrange the data into a matrix of users x movies, with the intersections representing an individual rating, then we can solve for the matrix of user-preferences and the matrix of movie characteristics, that when multiplied, give the completed matrix which we have just formed. Another way of explaining it is that we find similarities between movies and similarities between users. Then we can recommend movies that similar users liked, or we can recommend a movie that is similar to others that a user has liked.
The Recosystem package was built specifically to handle this type of matrix factorization. Recosystem takes three vectors: one for the users, one for the movies, and one for the ratings. We will include 2 approaches one without tuning the model parameters, and one after tuning the parameters. In tuning the parameters, multiple values were experimented with, however this process took a significant amount of computational time, so only these two models are included here.

#Results
###section that presents the modeling results and discusses the model performance

#Conclusion
###section that gives a brief summary of the report, its limitations and future work